\Large Carryover effects in free recall reveal how past experiences influence memories of future experiences

### 2023-01-16 paper/main.tex
added some notes on priming (discussion section)
added a discussion section on priming effects

In everyday experience, each moment's unique blend of contextual features
(where we are, who we are with, what else we are thinking of at the time, what
else we experience nearby in time, etc.) plays an important role in how we
interpret, experience, and remember that moment, and how we relate it to our
other experiences~\citep[e.g., for review see ][]{Mann20}. What are the
analogues of real-world contexts in laboratory tasks like the free recall
paradigm employed in our study? In general, modern formal accounts of free
recall~\citep{Kaha20} describe context as comprising a mix of (a) features
pertaining to or associated with each item and (b) other items and thoughts
experienced nearby in time, e.g., that might still be ``lingering'' in the
participant's thoughts at the time they study the item. Item features can
include semantic properties (i.e., features related to the item's meaning),
lexicographic properties (i.e., features related to the item's letters),
sensory properties (i.e., feature related to the item's appearance, sound,
smell, etc.), emotional proprties (i.e., features related to how meaningful the
item is, whether the item evokes positive or negative feelings, etc.),
utility-related properties (e.g., features that describe how an item might be
used or incorporated into a particular task or situation), and more.
Essentially any aspect of the participant's experience that can be
characterized, measured, or otherwise described can be considered to influence
the participant's mental context at the moment they experience that item.
Temporally proximal features include aspects of the participant's internal or
external experience that are \textit{not} specifically occuring at the moment
they encounter an item, but that nonetheless influence how they process the
item. Thoughts related to percepts, goals, expectations, other experiences, and
so on that might have been cued (directly or indirectly) by the participant's
recent experiences prior to the current moment all fall into this category.
Internally driven mental states, such as thinking about an experience unrelated
to the experiment, also fall into this category.

@@

>>>>

In real-world experience, each moment's unique blend of contextual features
(where we are, who we are with, what else we are thinking of at the time, what
else we experience nearby in time, etc.) plays an important role in how we
interpret, experience, and remember that moment, and how we relate it to our
other experiences~\citep[e.g., for review see ][]{Mann20}. What are the
analogues of real-world contexts in laboratory tasks like the free recall
paradigm employed in our study? In general, modern formal accounts of free
recall~\citep{Kaha20} describe context as comprising a mix of (a) features
pertaining to or associated with each item and (b) other items and thoughts
experienced nearby in time, e.g., that might still be ``lingering'' in the
participant's thoughts at the time they study the item. Item features can
include semantic properties (i.e., features related to the item's meaning),
lexicographic properties (i.e., features related to the item's letters),
sensory properties (i.e., feature related to the item's appearance, sound,
smell, etc.), emotional proprties (i.e., features related to how meaningful the
item is, whether the item evokes positive or negative feelings, etc.),
utility-related properties (e.g., features that describe how an item might be
used or incorporated into a particular task or situation), and more.
Essentially any aspect of the participant's experience that can be
characterized, measured, or otherwise described can be considered to influence
the participant's mental context at the moment they experience that item.
Temporally proximal features include aspects of the participant's internal or
external experience that are \textit{not} specifically occuring at the moment
they encounter an item, but that nonetheless influence how they process the
item. Thoughts related to percepts, goals, expectations, other experiences, and
so on that might have been cued (directly or indirectly) by the participant's
recent experiences prior to the current moment all fall into this category.
Internally driven mental states, such as thinking about an experience unrelated
to the experiment, also fall into this category.

@@

% explicit priming (resolve ambiguous stimuli, e.g., green eyes study)
% negative priming: supressing previously irrelevant stimuli or features hurts participants' recall for those stimuli when they're used on subsequent lists
% implicit priming (most like here): patterns in prior experiences sets expectations about what is to come, which can affect memory for upcoming stimuli
%    - participants in the reduced ({early, late}) conditions picked up on irrelevant features in early lists and when those features changed their memories for subsequent lists were affected.
%    - participants in the order manipulation conditions picked up on structure in temporal order on early lists, and that affected how they remembered later lists

When our ongoing experiences are ambiguous, we can draw on our past
experiences, expectations, and other real, perceived, or inferred cues to help
resolve the ambiguities. We may also be overtly or covertly ``primed'' to
influence how we are likely to resolve ambiuities. For example, before
listening to a story with several equally plausable interpretations, providing
participants with ``background'' information beforehand can lead them towards
one interpretation versus another~\citep{YeshEtal17}. More broadly, our
conscious and unconscious biases and preferences can influence not only how we
interpret high-level ambuities, but even how we process low-level sensory
information~\citep{KataEtal23}.

In more simplified scenarios, like list learning paradigms, the stimuli and
tasks participants encounter before studying a given list can influence what
and how they remember. For example, when participants are directed to supress,
disregard, or ignore ``distracting'' stimuli early on in an experiment,
participants often tend to remember those stimuli less well when they are
re-used as to-be-remembered targets later on in the experiment~\citep{Tipp85}.
In general, participants' memories can be influenced by a wide range of
positive and negative factors~\citep{BaloEtal92, ClayChat89, Donn88,
FlexTulv82, GottEtal12, HuanEtal04, Hube08, HubeEtal01, McNa94, Neel77, Rabi86,
TulvScha91, WatkEtal92, WiggMart98}.

The order manipulation conditions in our experiment show that participants can
also be primed to pick up on more subtle statistical structure in their
experiences, like the dynamics of how the presentation orders of stimuli vary
along particular feature dimensions. These order manipulations affected not
only how participants remembered the manipulated lists, but also how they
remembered \textit{future} lists with different (randomized) temporal
properties.
